### 2017-05-02 DDM_Report.Rmd
began ddm report

---
title: 'Closer look at DDM parameter estimates'
output:
github_document:
toc: yes
toc_float: yes
---

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(lme4)
library(GGally)
library(jsonlite)
library(psych)
library(rmarkdown)
library(psych)
library(stringr)
library(plotly)
library(jsonlite)
library(DT)
sem <- function(x) {sd(x, na.rm=T) / sqrt(length(x))}
render_this <- function(){rmarkdown::render('DDM_Report.Rmd', output_dir = '/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Report', html_notebook(toc = T, toc_float = T, code_folding = 'hide'))}
options(scipen = 1, digits = 4)
```

How much do hddm parameters change when fitting to whole vs partial sample

Summary of absolute difference in parameter estimates

```{r}
pre_correction <- read.csv('/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/retest_subs_test_data_preddm_correction.csv')
post_correction <- read.csv('/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/retest_subs_test_data.csv')

hddm_cols = grep('hddm', names(pre_correction), value =T)

pre_correction = pre_correction[,c("sub_id", hddm_cols)]
post_correction = post_correction[,c("sub_id", hddm_cols)]

pre_correction = pre_correction %>% mutate(pre_post = "pre")
post_correction = post_correction %>% mutate(pre_post = "post")

all_hddm = rbind(pre_correction, post_correction)

all_hddm %>%
  gather(key, value, -sub_id, -pre_post) %>%
  arrange(sub_id) %>%
  spread(pre_post, value) %>%
  mutate(abs_diff = pre-post) %>%
  group_by(key) %>%
  summarise(mean_abs_diff = mean(abs_diff, na.rm=T),
            median_abs_diff = median(abs_diff, na.rm=T),
            min_abs_diff = min(abs_diff, na.rm=T),
            max_abs_diff = max(abs_diff, na.rm=T),
            sd_abs_diff = sd(abs_diff, na.rm=T))
```

How to reliabilities change w sample size (refit from 50 to 150) both for EZ and hddm (gives a prescriptive recommendation on which method to use depending on sample size)

Turn off h of ddm and see if it makes a difference