### 2018-07-18 code/Boot_Figure.Rmd
changing strategy from using plotly for all measures to shiny app filtering tasks

---
title: "Bootstrapped Retest Reliability Figure"
output:
  html_document: default
  pdf_document: default
---

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(stringr)
library(plotly)
sem <- function(x) {sd(x, na.rm=T) / sqrt(length(x))}
options(scipen = 1, digits = 4, width = 500)
```